---
title: "MachineLearning"
author: "Fábio Franco Costa"
date: "April 24, 2015"
output: html_document
---


You should create a report describing  

* how you built your model,
* how you used cross validation  
* what you think the expected out of sample error is, 
* and why you made the choices you did.  
* You will also use your prediction model to predict 20 different test cases. 


```{r}
# Load packages and set seed
library("caret")
set.seed(666)

# Read Data
pmlData <- read.csv("pml-training.csv")
pmlFinalTest <- read.csv("pml-testing.csv") 

# Filter relevant columns

# Convert variables to factor 

# Preprocess training data

# Create trainning datasets
inTrain <- createDataPartition(y = pmlData$classe, p = 0.8, list = F)
trainning <- pmlData[inTrain,]
testing <- pmlData[-inTrain,]

# Fit first model: rpart
modelFit <- train(classe ~ .,
                  data = trainning,  
                  method=c("rpart"))

# Test the model
predictions <- predict(modelFit, newdata = testing)
confusionMatrix(predictions, testing$Survived)

```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
