---
title: "The Exponencial Distribution Study"
author: "DeltaSigma"
output: html_document
---
## Overview
In this brief study we will try to understand the behavior of the **Exponencial Distribution**. To do that, we will perform simulations and we will compare the results with the theoretical predictions. We hope that this study will be as useful to one who reads as it was useful to us that wrote it.

## Simulations
The first step will be to perform a 1000 simulation of exponentials and then generate 1000 averages of 40 exponentials. To do that we will use the following R code:

```{r}
# Run simulation of 1000 exponentials
nosim <- 1000
lambda <- 0.2
exp1000 <- rexp(nosim,lambda)

# Run 1000 simulation of average of 40 exponentials
mex = NULL
for (i in 1:nosim) mex = c(mex,mean(rexp(40,lambda)))
```

In our code **exp1000** hold the 1000 simulated exponentials and **mex** holds the 1000 averages of 40 exponentials.

## Sample Mean versus Theoretical Mean
Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers.

```{r, echo=FALSE}
# Set margins and multi column graph
library(ggplot2)
par(mar = rep(2, 4))
par(mfrow=c(1,2))

hist(exp1000)
hist(mex)

dat <- data.frame(
        x = c(exp1000,mex),
        t = factor(rep(c('exp','avgexp'), rep(nosim, 2)))
        )

g <- ggplot(dat, aes(x = x, fill = t)) + geom_histogram(binwidth=.3, colour = "black") 
g <- g + stat_function(fun = dnorm, size = 2)
g + facet_grid(. ~ t)



```

We can easily see that the exponential distribution does not look like a normal distribution but the distribution of averages of 40 exponentials does.

## Sample Variance versus Theoretical Variance
Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.

## Distribution
Via figures and text, explain how one can tell the distribution is approximately normal. 
