---
title: "The Exponencial Distribution Study"
author: "DeltaSigma"
output: html_document
---
## Overview
In this brief study we will try to understand the behavior of the **Exponencial Distribution**. To do that, we will perform simulations and we will compare the results with the theoretical predictions. We hope that this study will be as useful to one who reads as it was useful to us that wrote it.

## Simulations
The first step will be perform a 1000 simulation and generate an histogram to see how the distribution looks like. To to that we will use the following R code:

```{r}
# Set margins
par(mar = rep(2, 4))

# Run simulation of 1000 exponentials
hist(rexp(1000,lambda))

# Run 1000 simulation of average of 40 exponentials
mns = NULL
for (i in 1:1000) mns = c(mns,mean(rexp(40,lambda)))
hist(mns)
```

Include English explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes.

## Sample Mean versus Theoretical Mean
Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers.

## Sample Variance versus Theoretical Variance
Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.

## Distribution
Via figures and text, explain how one can tell the distribution is approximately normal. 



You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
