---
title: "Be Ready For Everything: An Analisys of NOAA Storm Database"
author: "Fábio Franco Costa"
date: "February 22, 2015"
output: html_document
---

    This document is a study of the most critical events in the last 10 years as recorded in NOAA database. Our objective is to create statistics to help the government to prepare to such situations. We selected the last 10 years because we found better and more comprehensive data in this period. Its our hope that you can use this information to save human lives.

# Data Processing

To perform this analisys we will load our data from the CSV file and than we will select only the data from the most recent ten years.

```{r Data Processing, cache=TRUE}
# Read file
file <- 'repdata_data_StormData.csv.bz2'
data <- read.csv(bzfile(file, 'repdata_data_StormData.csv'))

# Fix date format and get the last 5 years
data$BGN_DATE <- as.POSIXct(data$BGN_DATE, format="%m/%d/%Y %H:%M:%S")
data <- data[data$BGN_DATE > "2001-11-30", ]
```
We will select only the necessary fields in the columns to reduce our processing overhead.
```{r reduce columns}
# Keep only the columns needed
data <- data[,c("BGN_DATE","EVTYPE","FATALITIES","INJURIES",
                "PROPDMG","CROPDMG")]
```

Now we are going to remove the rows that have NAs for EVTYPE:

```{r remove NAs}
data <- data[!is.na(data$EVTYPE), ]
```

The next step is to summarize the data by **EVTYPE** and divide by 5 because we want to have an ideia of how frequent are this events on an yearly basis.
```{r Summary, cache=TRUE}
sumData = data.frame()
for (event in unique(data$EVTYPE)) {
        tDF <- data.frame("EVTYPE" = event,
                 "FATALITIES" = sum(data$FATALITIES[data$EVTYPE == event]),
                 "INJURIES" = sum(data$INJURIES[data$EVTYPE == event]),
                "PROPDMG" = sum(data$PROPDMG[data$EVTYPE == event]),
                "CROPDMG" = sum(data$CROPDMG[data$EVTYPE == event]))
        sumData <- rbind(sumData,tDF)
}
sumData$HEALTH <- sumData$FATALITIES + sumData$INJURIES
sumData$ECONOMIC <- sumData$PROPDMG + sumData$CROPDMG
sumData[2:6] <- sumData[2:6] / 10
```
Finaly, the TOP 5 events for public health are:
```{r health}
head(sumData[order(sumData$HEALTH, decreasing = T),],5)[1:3]
```

To get a sense of the event proportion for health you can look the following figure

```{r health graph, echo=FALSE}
# Plot
library(ggplot2)
h <- head(sumData[order(sumData$HEALTH, decreasing = T),],5)[,c(1,6)]
g <- ggplot(h, aes(x = h$EVTYPE, y = h$HEALTH))
g <- g + geom_bar(stat="identity")
g <- g + labs(x = "Events", y = "Casulties (Injuries + Fatalities")
g
```

The TOP 5 events for great economic consequences are:

```{r}
head(sumData[order(sumData$ECONOMIC, decreasing = T),],5)[,c(1,4,5)]
```

To get a sense of the event proportion for economic impact you can look the following figure

```{r economic graph, echo=FALSE}
# Plot
library(ggplot2)
h <- head(sumData[order(sumData$ECONOMIC, decreasing = T),],5)[,c(1,7)]
g <- ggplot(h, aes(x = h$EVTYPE, y = h$ECONOMIC))
g <- g + geom_bar(stat="identity")
g <- g + labs(x = "Events", y = "Economic Impact")
g
```

# Results  
Our conclusion is that the most important event for human health are **tornados**, and it is far worst than the second place. 

For economic impact, the government should be woried with the TOP 5. Their occurrence are responsable for a large amound of economic damage.

We hope that this report helps you prepare better for anything that may face you in the near future.